How to Not write Virtual Machines for Dynamic Languages

### 2007-05-27 talk/dyla2007/dyla.tex
Conclusion.

@@

\section{Conclusion}
XXX 
% ---- Bibliography ----
%\begin{small}
\bibliographystyle{abbrv}
\bibliography{dyla}
%\end{small}

>>>>

In the sequel, we will focus on the relative advantages and
inconvenients of the PyPy approach compared to the approach of
hand-writing a language implementation on top of an OO VM.

@@

\section{Conclusion}

Here are the two central points that we have asserted in the present
paper:

\begin{itemize}
\item \emph{Do not write dynamic language implementations ``by hand''.}
Writing them more abstractly, at a higher level, has mostly only
advantages, among them avoiding a proliferation of implementations
growing out of sync.  Writing interpreters both flexibly and efficiently
is difficult, and meta-programming is a good way to achieve it.
Moreover, this is not incompatible with targetting and benefiting from
existing good object-oriented virtual machines like the Java and .NET
ones.

\item \emph{Do not write VMs ``by hand''.}
Writing language-specific virtual machines is a time-consuming task for
medium to large languages.  Unless large amounts of resources can be
invested, the resulting VMs are bound to have limitations which lead to
the emergence of many implementations, a fact that is taxing precisely
for a community with limited resources.  (This is of course even more
true for general-purpose VMs.)
\end{itemize}

As a better alternative, we advocate a more general usage of
meta-programming:

\begin{itemize}
\item \emph{Let's write more meta-programming translation toolchains.}
Aside from the advantages described in section
\ref{sect:metaprogramming}, a translation toolchain need not be
standardized for inter-operability, but can be tailored to the needs of
each project.  Diversity is good; there is no need to attempt to
standardize on a single OO VM.
\end{itemize}

The approach we outlined is actually just one in a very large, mostly
unexplored design space; it is likely that some of the choices made in
PyPy will turn out to be suboptimal.  We are hoping that other
toolchains will emerge over time, exploring other aspects and proposing
other solutions.  By their ``meta'' nature, these multiple approachs
should be easier to bridge together than, say, multiple OO VMs with
different object and runtime models.  We believe that further research
in this area might open the door to better solutions for
interoperability in general -- e.g. high-level bridges instead of
(virtual-)machine-level ones, enabled by cross-translation.

We believe this to be ultimately a better investment of efforts than the
development of more advanced general-purpose OO VMs.

% ---- Bibliography ----
%\begin{small}
\bibliographystyle{abbrv}
\bibliography{dyla}
%\end{small}
